---
title: "General Social Survey"
author: "Yolanda Landeros"
date: "Year 2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"

mydata5 <- filter(mydata4, CAPPUN=="FAVOR" | CAPPUN=="OPPOSE")

```


# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. First explain what you make out of comparing how people voted in 2012 with 2016.

How are the "voted in 2016" and "voted in 2012" related?  
$H_0:$ voting in 2016 and 2012 are not related.  



## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
 use categorical variables, and a chi-squared test to test the hypothesis that voting in 2016 is not independent of voting in 2012

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->


### Descriptive Results

we use a barchart to show the voting trends between 2012 and 2016

```{r}
barchartGC(~VOTE16+VOTE12, data=mydata4, type = "percent" ,xlab= 'voted in 2016') 
chisqtestGC(~VOTE16+VOTE12, data=mydata4)
votetab=xtabs(~VOTE16+VOTE12, data=mydata4)
fisher.test(votetab)
```

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~VOTE16+VOTE12, data=mydata4, type = "percent" ,xlab= 'voted in 2016') 
plot(xtabs(~VOTE16+VOTE12, data=mydata4))
plot(votetab)
summary(xtabs(~VOTE16+VOTE12, data=mydata4))
```


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!--> 
```{r}
xtabs(~VOTE16+VOTE12, data=mydata4)
```


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(~VOTE16+VOTE12, data=mydata4)
votetab=xtabs(~VOTE16+VOTE12, data=mydata4)
fisher.test(votetab)
```

$H_a$: if you voted in 2012 it is likely you would vote in 2016

we reject the null hypothesis, we have evidence that voting in 2012 is related to voting in 2016. If you voted in 2012 it is likely you will vote in 2016.

# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Now answer Is belief in the bigbang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)

$H_0:$ believing in the bing bang is not independent from the belief in postlife.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

 use categorical variables, and a chi-squared test to test the hypothesis that the belief of the bigbang is not dependent of afterlife.


##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results
it shows the results between bigbang theory and post life beliefs.

```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata3, type = "percent" , xlab= 'believe in postlife' , main= 'believe in bigbang') 
chisqtestGC(~BIGBANG+POSTLIFE, data=mydata3)
BELIEFS=xtabs(~BIGBANG+POSTLIFE, data=mydata3)
fisher.test(BELIEFS)
```


#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata3, type = "percent" , xlab= 'believe in postlife' , main= 'believe in bigbang')
plot(xtabs(~BIGBANG+POSTLIFE, data=mydata3))
summary(xtabs(~BIGBANG+POSTLIFE, data=mydata3))
```


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
xtabs(~BIGBANG+POSTLIFE, data=mydata3)
```


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

$H_a$: the null hypothesis is people who believe in bigbang do not believe in the after life.

we aren't rejecting the null hypothesis, there is no evidence.


```{r}
chisqtestGC(~BIGBANG+POSTLIFE, data=mydata4)
BELIEFS=xtabs(~BIGBANG+POSTLIFE, data=mydata4)
fisher.test(BELIEFS)
```


# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This time compare the opinion on death penalty (CAPPUN) with opinion on courts (COURTS).

$H_0:$ the opinion with death penalty is not independet with the opinion with the opinion with the courts.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
 use categorical variables, and a chi-squared test to test the hypothesis that the opinions of the death penalty is not dependent with opions of the court.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results
the graph shows the results between death penalty and courts.
```{r}
barchartGC(~CAPPUN+COURTS, data=mydata5, type = "percent" , xlab= 'view on CAPPUN' , main= 'view on the COURTS') 
chisqtestGC(~CAPPUN+COURTS, data=mydata5)
OPINION=xtabs(~CAPPUN+COURTS, data=mydata5)
fisher.test(OPINION)
```

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
barchartGC(~CAPPUN+COURTS, data=mydata5, type = "percent" , xlab= 'view on CAPPUN' , main= 'view on the COURTS') 
chisqtestGC(~CAPPUN+COURTS, data=mydata5)
plot(xtabs(~CAPPUN+COURTS, data=mydata5))
summary(xtabs(~CAPPUN+COURTS, data=mydata5))
```


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
xtabs(~CAPPUN+COURTS, data=mydata5)
```


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
$H_a$: the null hypothesis is there is no relation between opinions between death penalty and the op8inion of the court.

we reject the null hypothesis because we do find evidence suggestion there is a relation.



```{r}
chisqtestGC(~CAPPUN+COURTS, data=mydata5)
OPINION=xtabs(~CAPPUN+COURTS, data=mydata5)
fisher.test(OPINION)
```



# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
The hypothesis was that voting in 2012 is related. We used a chi-squared test if independace the p value is equal to 0. Therefore, the null hypothesis is rejected. We reject the null hypothesis because the p valuie is less than .35, meaning the null hypothesis is related. we reject the null since we have evidence that the belief in death penalty is related to the opinion to the court system and if youn do not believe in the edeath penalty it is likely you did n ot believe in the court system.
